---
title: "BIS 537 Final Project"
author: "Ruyi Liu, Jingchen Liang"
date: '2022-12-01'
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r,warning=FALSE,message=FALSE}
library(tidyverse)
library(corrplot)
library(MASS)
library(PSweight)
library(VGAM)
library(brglm2)
library(mvtnorm)
library(randomForest)
library(DMwR)
library(latex2exp)
```


# Part 0. Data Import

```{r, warning=FALSE, message=FALSE}
# load data
# data link: https://www.kaggle.com/datasets/kartik2khandelwal/no-title
liver <- read.csv("liver.csv",na.strings = c("", "NA"))
dim(liver)  ## 6800 obs & 20 cols
colnames(liver)
```

# Part 1. Data Preprocessing

```{r}
# delete all observations when drug (Z) is missing
liver1 <- liver[which(!is.na(liver$Drug)),]

# each patient only has one record
length(unique(liver1$ID))  

colSums(is.na(liver1))
# select variables
liver2 <- liver1[,c("Drug","Age","Sex","Hepatomegaly","Edema","Bilirubin","Albumin","SGOT","Platelets","Prothrombin","Stage")]
liver3 <- liver2[complete.cases(liver2), ]   # 1995 obs & 11 cols
dim(liver3)
table(liver3$Stage)

# reformat the variables
liver3$Drug <- ifelse(liver3$Drug == "D-penicillamine", 1, 0)   # D-penicillamine = 1, placebo = 0
liver3$Sex <- ifelse(liver3$Sex == "F", 1, 0)  # F = 1, M = 0
liver3$Hepatomegaly <- ifelse(liver3$Hepatomegaly == "Y", 1, 0)  # Y = 1, N = 0
liver3$Edema <- ifelse(liver3$Edema == "N", 1, 0)  # Y = 1, N = 0
liver3$Age <- ceiling(liver3$Age/365)
```


```{r}
liver_final <- liver3
# check correlations
liver_Cor <- cor(liver_final[,c(2,6:10)])
##Visualize correlations
corrplot.mixed(liver_Cor, tl.col="black", tl.pos = "lt")    # correlations are low
```


# Part 2. Propensity Score Analysis

__1. Direct regression estimator__ 

We use random forest algorithm to find the direct regression estimator. By definition, the regression estimator for ATE is 
$$\hat{\tau}^{ATE}= N^{-1}\sum_{i=1}^{N}(\hat{m}_1(X_i)-\hat{m}_0(X_i))$$ 

```{r,warning=FALSE,message=FALSE}
## direct regression estimator
# use random forest
set.seed(60)
liver_final1 <- liver_final
liver_final1$Stage_binary <- ifelse(liver_final$Stage == 4, 1, 0)
liver_final1$Stage <- NULL
liver_final_1 <- subset(liver_final1, liver_final1$Drug == 1)
liver_final_0 <- subset(liver_final1, liver_final1$Drug == 0)
# regression among the untreated Z = 0
model_a0 <- glm(as.factor(Stage_binary) ~., family = binomial,data = liver_final_0)
mu_hat_a0 <- predict(model_a0,newdata=liver_final1,type="response")
# regression among the untreated Z = 1
model_a1 <- glm(as.factor(Stage_binary) ~., family = binomial,data = liver_final_1)
mu_hat_a1 <- predict(model_a1,newdata=liver_final1,type="response")

# estimate the ATE
ATE_direct <- sum(as.numeric(mu_hat_a1)-as.numeric(mu_hat_a0))/nrow(liver_final1)
ATE_direct

# estimate the risk ratio
RR_direct <- mean(mu_hat_a1)/mean(mu_hat_a0)
RR_direct
```



```{r,warning=FALSE,message=FALSE}
# use bootstrap to find the standard error
set.seed(60)
n <- nrow(liver_final)
# perform 1000 bootstrap simulations
B <- 1000
idx <- 1:nrow(liver_final)
ATT_direct_estimates <- c()
RR_direct_estimates <- c()
for(i in 1:B){
  sample_index <- sample(idx, n, replace=TRUE)
  new_sample <- liver_final[sample_index,]
  new_sample_1 <- subset(liver_final1, liver_final1$Drug == 1)
  new_sample_0 <- subset(liver_final1, liver_final1$Drug == 0)
  # regression among the untreated Z = 0
  model_a0 <- glm(as.factor(Stage_binary) ~., family = binomial,data = new_sample_0)
  mu_hat_a0 <- predict(model_a0,newdata=new_sample,type="response")
  # regression among the untreated Z = 1
  model_a1 <- glm(as.factor(Stage_binary) ~., family = binomial,data = new_sample_1)
  mu_hat_a1 <- predict(model_a1,newdata=new_sample,type="response")
  ATT_direct_estimates[i] <- sum(as.numeric(mu_hat_a1)-as.numeric(mu_hat_a0))/nrow(new_sample)
  RR_direct_estimates[i] <- mean(mu_hat_a1)/mean(mu_hat_a0)
}
sd(ATT_direct_estimates)
sd(RR_direct_estimates)
```

The estimated ATE using direct regression estimator is 0.008289998, and its standard error is 0.001075567.  
The estimated risk ratio using direct regression estimator is 1.01624, and its standard error is 0.00212431.    


__2. IPW with normalized weights__ 


```{r}
liver_final1 <- liver_final
liver_final1$Stage_binary <- ifelse(liver_final$Stage == 4, 1, 0)
liver_final1$Stage <- NULL
# run a logistic regression model with all pre-treatment variable
logit_model1 = glm(Drug ~. - Stage_binary, family = "binomial",data = liver_final1)
# get the propensity score
propensity_score1 = predict(logit_model1, type = "response")
result1 = data.frame(liver_final1,propensity_score=propensity_score1)
# find inverse probability weighting (IPW) for Z = 1 (W1) and Z = 0 (W0)
result1$IPW = ifelse(result1$Drug == 0, 1/(1-propensity_score1), 1/propensity_score1)
head(result1)
```

```{r}
# subset the result data
# for Z = 0
result1_Z0 = subset(result1,result1$Drug == 0)
# for Z = 1
result1_Z1 = subset(result1,result1$Drug == 1)
# get the unweighted and weighted variance for each variable
s0 = apply(result1_Z0[2:12], 2, var)
s1 = apply(result1_Z1[2:12], 2, var)
# calculate the absolute standardized difference (ASD) for each variable
pre_var_list = colnames(result1)[2:12]
ASD = c()
for(i in 1:length(pre_var_list)){
 ASD[i] = abs(sum(result1_Z1$IPW*result1_Z1$Drug*result1_Z1[pre_var_list[i]])/sum(result1_Z1$IPW*result1_Z1$Drug) -
 sum(result1_Z0$IPW*(1-result1_Z0$Drug)*result1_Z0[pre_var_list[i]])/sum(result1_Z0$IPW*(1-result1_Z0$Drug)))/
 sqrt((s0[pre_var_list[i]]+s1[pre_var_list[i]])/2)
}
# show results
ASD_res = data.frame(pre_var_list,ASD)
ASD_res
```

```{r}
# plot the histogram
ggplot(result1, aes(x = propensity_score)) +
 geom_histogram(aes(color = factor(Drug), fill = factor(Drug)),
 position = "identity", bins = 30, alpha = 0.3) +
 labs(title="Overlaying Histogram of the Estimated Propensity Scores by Treatment Groups",
 color="Drug",fill="Drug",x="propensity score") + theme_bw()
```

```{r, fig.width=12,fig.height=11.5, fig.asp = .5}
# plot the loveplot for the weighted ASD
summary_PS <- SumStat(ps.formula = Drug ~ Age+Sex+Hepatomegaly+Edema+Bilirubin+Albumin+SGOT+Platelets+Prothrombin, 
                      data = liver_final1, weight = "IPW")
plot(summary_PS, type = "balance", metric = "ASD",threshold = 0.1) # based on ASD
```

The histogram shows a satisfactory overlap of propensity scores between treatment groups.From the love plot, we find that the unweighted mean differences of the original ASD are larger than the values of IPW. Of these three methods, the IPW keeps each mean difference below 0.1, and the ASD of each covariate is close to 0, which provides the best balance.



Given $w_1(X_i)=\frac{h(X_i)}{e(X_i)}$, and $w_0(X_i)=\frac{h(X_i)}{1-e(X_i)}$, and by the definition of IPW, since $w_1(X_i)=\frac{1}{e(X_i)}$, and $w_0(X_i)=\frac{1}{1-e(X_i)}$, so $h(X_i) = 1$ 


```{r}
# Estimate the ATE
# get w1 and w0
w1 = 1/result1$propensity_score
w0 = 1/(1-result1$propensity_score)
# find the ATE for using non-normalized weights
ATT_IPW_normalized = sum(w1*result1$Drug*result1$Stage_binary)/sum(w1*result1$Drug)-
 sum(w0*(1-result1$Drug)*result1$Stage_binary)/sum(w0*(1-result1$Drug))
ATT_IPW_normalized
RR_IPW_normalized <- (sum(w1*result1$Drug*result1$Stage_binary)/sum(w1*result1$Drug))/(sum(w0*(1-result1$Drug)*result1$Stage_binary)/sum(w0*(1-result1$Drug)))
RR_IPW_normalized
```


```{r}
# use bootstrap to find the standard error
set.seed(100)
n = nrow(liver_final1)
# perform 1000 bootstrap simulations
B = 1000
idx = 1:nrow(liver_final1)
ATT_estimate_IPW = c()
RR_estimate_IPW = c()
for(i in 1:B){
 sample_index = sample(idx, n, replace=TRUE)
 new_sample = liver_final1[sample_index,]
 model = glm(Drug ~. - Stage_binary, family = "binomial",data = new_sample)
 new_sample$propensity_score = predict(model,data=new_sample,type = "response")
 # get w1 and w0
 w1 = 1/new_sample$propensity_score
 w0 = 1/(1-new_sample$propensity_score)
 ATT_estimate_IPW[i] = sum(w1*new_sample$Drug*new_sample$Stage_binary)/sum(w1*new_sample$Drug)-
 sum(w0*(1-new_sample$Drug)*new_sample$Stage_binary)/sum(w0*(1-new_sample$Drug))
 RR_estimate_IPW[i] = (sum(w1*new_sample$Drug*new_sample$Stage_binary)/sum(w1*new_sample$Drug))/(sum(w0*(1-new_sample$Drug)*new_sample$Stage_binary)/sum(w0*(1-new_sample$Drug)))
}
sd(ATT_estimate_IPW)
sd(RR_estimate_IPW)
```


The estimated ATE using IPW estimate using normalized weights is 0.008676144, and its standard error is 0.02264897.  
The estimated risk ratio using IPW estimate using normalized weights is 1.016999, and its standard error is 0.04505959.      




__3. Doubly-Robust estimator__ 

According to the notes in Lecture 7, we have doubly robust estimator is:  

$$
\begin{aligned}
\hat {\tau}_{dr} = \frac{1}{N}\sum_{i=1}^{N} \left[\hat{m}_1(X_i)+\frac{Z_i(Y_i-\hat{m_1}(X_i))}{\hat{e}(X_i)} \right] -\frac{1}{N}\sum_{i=1}^{N} \left[\hat{m}_0(X_i)+\frac{(1-Z_i)(Y_i-\hat{m_0}(X_i))}{1-\hat{e}(X_i)} \right]
\end{aligned}
$$

```{r,warning=FALSE,message=FALSE}
# first run a logistic regression model to get the propensity score
model_DR = glm(Drug ~. - Stage_binary, family = "binomial",data = liver_final1)
# get the propensity score
propensity_score_DR = predict(model_DR, type = "response")
# subset the data by treatment
liver_final_1 = subset(liver_final1, liver_final1$Drug == 1)
liver_final_0 = subset(liver_final1, liver_final1$Drug == 0)
# fit regression models based on the observed outcome in each treatment group
model_1 = glm(Stage_binary ~., family = "binomial",data = liver_final_1)
model_0 = glm(Stage_binary ~., family = "binomial",data = liver_final_0)
# get outcome prediction for Z = 1
prediction_1 = predict(model_1, type="response",newdata=liver_final1)
# get outcome prediction for Z = 0
prediction_0 = predict(model_0, type="response",newdata=liver_final1)
# find the ATE - risk difference
ATE_DR = sum(prediction_1+liver_final1$Drug*(liver_final1$Stage_binary-prediction_1)/propensity_score_DR)/nrow(liver_final1)-
      sum(prediction_0+(1-liver_final1$Drug)*(liver_final1$Stage_binary-prediction_0)/(1-propensity_score_DR))/nrow(liver_final1)
ATE_DR
RR_DR = (sum(prediction_1+liver_final1$Drug*(liver_final1$Stage_binary-prediction_1)/propensity_score_DR)/nrow(liver_final1))/(sum(prediction_0+(1-liver_final1$Drug)*(liver_final1$Stage_binary-prediction_0)/(1-propensity_score_DR))/nrow(liver_final1))
RR_DR
```

```{r,warning=FALSE,message=FALSE}
# use bootstrap to find the standard error for RD and RR using IPW
set.seed(100)
n = nrow(liver_final1)
# perform 1000 bootstrap simulations
B = 1000
idx = 1:nrow(liver_final1)
ATE_boot_IPW = c()
RR_boot_IPW = c()
for(i in 1:B){
  sample_index = sample(idx, n, replace=TRUE)
  new_sample = liver_final1[sample_index,]
  model_g = glm(Drug ~. - Stage_binary, family = "binomial",data = new_sample)
  # get the propensity score
  propensity_score_DR = predict(model_DR, type = "response")
  # subset the data by treatment
  liver_final_1 = subset(new_sample, new_sample$Drug == 1)
  liver_final_0 = subset(new_sample, new_sample$Drug == 0)
  # fit regression models based on the observed outcome in each treatment group
  model_1 = glm(Stage_binary ~., family = "binomial",data = liver_final_1)
  model_0 = glm(Stage_binary ~., family = "binomial",data = liver_final_0)
  # get outcome prediction for Z = 1
  prediction_1 = predict(model_1, type="response",newdata=new_sample)
  # get outcome prediction for Z = 0
  prediction_0 = predict(model_0, type="response",newdata=new_sample)
  # find the ATE - risk difference
  ATE_boot_IPW[i] = sum(prediction_1+new_sample$Drug*(new_sample$Stage_binary-prediction_1)/propensity_score_DR)/nrow(new_sample)-
        sum(prediction_0+(1-new_sample$Drug)*(new_sample$Stage_binary-prediction_0)/(1-propensity_score_DR))/nrow(new_sample)
  RR_boot_IPW[i] = (sum(prediction_1+new_sample$Drug*(new_sample$Stage_binary-prediction_1)/propensity_score_DR)/nrow(new_sample))/(sum(prediction_0+(1-new_sample$Drug)*(new_sample$Stage_binary-prediction_0)/(1-propensity_score_DR))/nrow(new_sample))
}
sd(ATE_boot_IPW)
sd(RR_boot_IPW)
```


The estimated ATE using doubly-Robust estimator is 0.008913068, and its standard error is 0.02284838.  
The estimated risk ratio using doubly-Robust estimator is 1.017467, and its standard error is 0.04544266.        

```{r}
# sensitivity analysis
1.01747 - qnorm(0.975)*0.04544266
1.01747 + qnorm(0.975)*0.04544266
```



__4. Propensity score stratification__ 


```{r}
logit_model2 = glm(Drug ~. - Stage_binary, family = "binomial",data = liver_final1)
# get the propensity score
propensity_score2 = predict(logit_model2, type = "response",liver_final1)
result2 = data.frame(liver_final1,propensity_score = propensity_score2)
# stratification method - cut the propensity scores into 5 strata
strata = cut(result2$propensity_score,breaks = quantile(c(0,result2$propensity_score), prob = seq(0,1,1/5),include.lowest = TRUE))
# relevel the strata
levels(strata) = c(1,2,3,4,5)
Stage_4 <- liver_final$Stage
result2 = data.frame(result2,strata,Stage_4)
#p1_binary = c()
#p2_binary = c()
#OR_binary = c()
# smaller than or equal to
p11 <- c()
p10 <- c()
p21 <- c()
p20 <- c()
p31 <- c()
p30 <- c()

OR_1 <- c()
OR_2 <- c()
OR_3 <- c()
# odds ratio pro cumulative prob
for(i in 1:5){
 strata_data = result2[result2$strata==i,]
 p11[i] = nrow(strata_data[strata_data$Drug==1 & strata_data$Stage_4<=1,])/length(which(strata_data$Drug==1))
 p10[i] = nrow(strata_data[strata_data$Drug==0 & strata_data$Stage_4<=1,])/length(which(strata_data$Drug==0))
 p21[i] = nrow(strata_data[strata_data$Drug==1 & strata_data$Stage_4<=2,])/length(which(strata_data$Drug==1))
 p20[i] = nrow(strata_data[strata_data$Drug==0 & strata_data$Stage_4<=2,])/length(which(strata_data$Drug==0))
 p31[i] = nrow(strata_data[strata_data$Drug==1 & strata_data$Stage_4<=3,])/length(which(strata_data$Drug==1))
 p30[i] = nrow(strata_data[strata_data$Drug==0 & strata_data$Stage_4<=3,])/length(which(strata_data$Drug==0))
}

OR_1 = (mean(p11)*(1-mean(p10)))/(mean(p10)*(1-mean(p11)))
OR_1
OR_2 = (mean(p21)*(1-mean(p20)))/(mean(p20)*(1-mean(p21)))
OR_2
OR_3 = (mean(p31)*(1-mean(p30)))/(mean(p30)*(1-mean(p31)))
OR_3
```


# Part 3. Ordinal Logistic Regression

```{r, warning=FALSE, message=FALSE}
## fit ordered logit model
# specify Hess=TRUE to have the model return the observed information matrix from optimization (called the Hessian) which is used to get standard errors.
liver_final_ordinal <- liver_final
liver_final_ordinal$Stage <- as.factor(liver_final_ordinal$Stage)
levels(liver_final_ordinal$Stage) <- c("inflammation","fibrosis","cirrhosis","liver_failure")
ordinal_reg <- polr(Stage ~ ., data = liver_final_ordinal, Hess=TRUE, method="probit")
summary(ordinal_reg)
table(predict(ordinal_reg))

# find the p-values
ctable <- coef(summary(ordinal_reg))
p <- round(pnorm(abs(ctable[, "t value"]), lower.tail = FALSE) * 2,4)
ctable <- cbind(ctable, "p value" = p)
ctable
# find the confidence interval
ci <- confint(ordinal_reg)

# find the odds ratios
exp(coef(ordinal_reg))

# odds ratios and CI
exp(cbind(OR = coef(ordinal_reg), ci))
```


For patients whose use drug, the odds of being more likely to higher stage (i.e., cirrhosis or fibrosis versus inflammation) is 1.067 times that of patients using placebo, holding constant all other variables (not effective).


```{r, warning=FALSE, message=FALSE}
## fit ordered logit model if we scale the numeric variables - we obtain the same answers
# specify Hess=TRUE to have the model return the observed information matrix from optimization (called the Hessian) which is used to get standard errors.
liver_final_ordinal <- liver_final
scale_liver <- data.frame(cbind(liver_final_ordinal[,c(1,3,4,5,11)],scale(as.matrix(liver_final_ordinal[,c(2,6:10)]))))
head(scale_liver)

scale_liver$Stage <- as.factor(scale_liver$Stage)
levels(scale_liver$Stage) <- c("inflammation","fibrosis","cirrhosis","liver_failure")
ordinal_reg <- polr(Stage ~ ., data = scale_liver, Hess=TRUE)
summary(ordinal_reg)

# find the p-values
ctable <- coef(summary(ordinal_reg))
p <- round(pnorm(abs(ctable[, "t value"]), lower.tail = FALSE) * 2,4)
ctable <- cbind(ctable, "p value" = p)
ctable
# find the confidence interval
ci <- confint(ordinal_reg)
ci
```


```{r}
# Y <= 1 | x = 1 / Y > 1 | x = 1
exp(-2.6396)/exp(0.0650427)
# Y <= 1 | x = 0 / Y > 1 | x = 0
exp(-2.6396)
# Y <= 2 | x = 1 / Y > 2 | x = 1
exp(-0.9371)/exp(0.0650427)
# Y <= 2 | x = 0 / Y > 2 | x = 0
exp(-0.9371)
# Y <= 3 | x = 1 / Y > 3 | x = 1
exp(-0.1110)/exp(0.0650427)
# Y <= 3 | x = 0 / Y > 3 | x = 0
exp(-0.1110)

# find odds ratio
# 1 vs 2,3,4
exp(-0.0650427)
# 1,2 vs 3,4
exp(-0.0650427)
# 1,2,3 vs 4
exp(-0.0650427)
```


# Part 4. Bayesian inference using Gibbs sampling

```{r}
a <- liver_final
a1 <- subset(a,a$Stage == 3 | a$Stage == 4)
a1$Sex <- as.factor(a1$Sex)
a1$Hepatomegaly <- as.factor(a1$Hepatomegaly)
a1$Edema <- as.factor(a1$Edema)
a1$Stage <- as.factor(a1$Stage)
new_df1 <- SMOTE(Stage ~., data = a1, perc.over = 600, perc.under = 100)
a2 <- subset(a,a$Stage == 1 | a$Stage == 2)
a2$Sex <- as.factor(a2$Sex)
a2$Hepatomegaly <- as.factor(a2$Hepatomegaly)
a2$Edema <- as.factor(a2$Edema)
a2$Stage <- as.factor(a2$Stage)
new_df2 <- SMOTE(Stage ~., data = a2, perc.over = 1600, perc.under = 100)
new_df <- as.data.frame(rbind(new_df2,new_df1))
liver_final_ordinal <- new_df
table(liver_final_ordinal$Stage)
liver_final_ordinal$Stage <- as.factor(liver_final_ordinal$Stage)
levels(liver_final_ordinal$Stage) <- c("inflammation","fibrosis","cirrhosis","liver_failure")
ordinal_reg <- polr(Stage ~ ., data = liver_final_ordinal, Hess=TRUE, method="logistic")
summary(ordinal_reg)
table(predict(ordinal_reg))
```


```{r}
set.seed(60)
new_df$Sex <- as.numeric(new_df$Sex)
new_df$Hepatomegaly <- as.numeric(new_df$Hepatomegaly)
new_df$Edema <- as.numeric(new_df$Edema)
new_df$Stage <- as.numeric(new_df$Stage)

scale_liver <- as.data.frame(scale(as.matrix(new_df[,c(2:10)])))
scale_liver$Drug <- new_df$Drug
scale_liver$Stage <- new_df$Stage
n <- nrow(scale_liver)
mc_num <- 10000
X <- as.matrix(scale_liver[,1:10])
beta <- matrix(0,nrow=mc_num,ncol=10)
Z <- matrix(0,nrow=mc_num,ncol=n)
Z_miss <- matrix(0,nrow=mc_num,ncol=n)
Y_miss <- matrix(0,nrow=mc_num,ncol=n)
test_stat1 <- rep(0,mc_num)
test_stat2 <- matrix(0,nrow=mc_num,ncol=4)
test_stat3 <- matrix(0,nrow=mc_num,ncol=4)

# initialize values
beta[1,] <- rep(2,10)
Z[1,] <- rep(10,n)
Z_miss[1,] <- rep(1,n)
Y_miss[1,] <- rep(1,n)
test_stat1[1] <- 1
test_stat2[1,] <- rep(0.5,4)
test_stat3[1,] <- rep(0.5,4)

Z_miss_data <- scale_liver
Z_miss_data$Drug <- ifelse(Z_miss_data$Drug == 0, 1, 0)
Z_miss_data$Stage <- as.factor(Z_miss_data$Stage)
for(i in 2:mc_num){
  # update beta
  beta[i,] <- rmvnorm(n=1,mean=(n/(n+1))*solve(t(X)%*%X)%*%t(X)%*%Z[i-1,],sigma=(n/(n+1))*solve(t(X)%*%X))
  # update Z
  for(j in 1:n){
    # delta
    Y_idx <- scale_liver[j,]$Stage
    rest <- scale_liver[-j,]
    maximum <- max(Z[i-1,][which(rest$Stage <= Y_idx)])
    minimum <- min(Z[i-1,][which(rest$Stage >= Y_idx)])
    delta <- ifelse(Z[i-1,j] >= maximum & Z[i-1,j] <= minimum, 1, 0)
    Z[i,j] <- rnorm(n=1,mean=X[j,]%*%beta[i,],sd=1)*delta
  }
  # update Z_miss
  for(k in 1:n){
    Z_miss[i,k] <- as.matrix(Z_miss_data[,1:10][k,])%*%beta[i,] + rnorm(1,0,1)
  }
  # update Y_miss
  X_beta <- c(as.matrix(scale_liver[,1:10])%*%beta[i,])
  model <- polr(as.factor(scale_liver$Stage) ~ as.matrix(scale_liver[,1:10])%*%beta[i,], method = "probit")
  Y_miss[i,] <- predict(model, newdata=Z_miss[i,])
  
  # update test statistics1
  df_YZ <- as.data.frame(cbind(Y_obs = scale_liver$Stage, Y_mis = Y_miss[i,], Z_obs = scale_liver$Drug, Z_mis = 1-scale_liver$Drug))
  Y_1 <- ifelse(df_YZ$Z_obs == 1,df_YZ$Y_obs,df_YZ$Y_mis)
  Y_0 <- ifelse(df_YZ$Z_obs == 0,df_YZ$Y_obs,df_YZ$Y_mis)
  df_potential_outcome <- as.data.frame(cbind(Y_1,Y_0))
  mat_P <- matrix(0,4,4)
  for(m in 1:4){
    for (n in 1:4){
      mat_P[m,n] <- nrow(df_potential_outcome[df_potential_outcome$Y_0==m & df_potential_outcome$Y_1==n,])/nrow(df_potential_outcome)
    }
  }
  P_0_rowsum <- rowSums(mat_P)
  P_1_colsum <- colSums(mat_P)
  diff_mat_v <- as.matrix(P_1_colsum-P_0_rowsum)
  test_stat1[i] <- norm(diff_mat_v, type="1")
  
  # update test statistics2 - distributional causal effect
  for(u in 1:4){
    test_stat2[i,u] <- nrow(df_potential_outcome[df_potential_outcome$Y_1 >= u,])/nrow(df_potential_outcome) - 
      nrow(df_potential_outcome[df_potential_outcome$Y_0 >= u,])/nrow(df_potential_outcome)
  }
  # update test statistics3 - M-test
  for(v in 1:4){
    test_stat3[i,v] <- median(df_potential_outcome$Y_1[which(df_potential_outcome$Y_0 == v)])
  }
}
```


```{r}
burnin <- 2000
est1 <- mean(test_stat1[(burnin+1):10000],na.rm = T)
sd1 <- sd(test_stat1[(burnin+1):10000],na.rm = T)
quantile1 <- quantile(test_stat1[(burnin+1):10000],probs=c(0.025,0.5,0.975),na.rm = T)
res_table1 <- as.data.frame(cbind(est1,sd1,quantile1[1],quantile1[2],quantile1[3]))
colnames(res_table1) <- c("est","sd","lower","median","upper")

est2 <- colMeans(test_stat2[(burnin+1):10000,])
sd2 <- apply(test_stat2[(burnin+1):10000,],2,sd)
quantile2 <- t(apply(test_stat2[(burnin+1):10000,],2,quantile, probs=c(0.025,0.5,0.975)))
res_table2 <- as.data.frame(cbind(est2,sd2,quantile2))
colnames(res_table2) <- c("est","sd","lower","median","upper")
# create the table
res_table <- rbind(res_table1,res_table2)
rownames(res_table) <- c("Test_stat1","Test_stat2_1","Test_stat2_2","Test_stat2_3","Test_stat2_4")
res_table
```


```{r}
bayesian_res <- as.data.frame(cbind(test_stat1[(burnin+1):10000],test_stat2[(burnin+1):10000,]))
colnames(bayesian_res) <- c("test_stat1","j = 1","j = 2","j = 3","j = 4")
ggplot(bayesian_res, aes(x=test_stat1)) + geom_histogram(bins=30, color="black",position = "identity") +
  labs(title="Posterior Distribution of One-Dimensional Estimand") + xlab(TeX("$\\d(P^{(0)},P^{(1)})$")) + theme_bw()
```


```{r}
bayesian_res[,3:5] %>% pivot_longer(cols = c("j = 2","j = 3","j = 4"), names_to = "delta", values_to = "values") %>%
  ggplot(aes(x = values, colour= factor(delta), fill = factor(delta)))+geom_histogram(position = "identity",alpha = 0.3,bins=30,) +  
  labs(title=TeX("$\\Posterior\\ Distribution \\ of\\  Delta_j$"),color=TeX("$\\Delta_j$"),fill=TeX("$\\Delta_j$"),x="propensity score") +
  xlab(TeX("$\\Delta_j$")) + scale_fill_manual(values = c("#FF6666", "#00AFBB", "#E69F00")) + 
  scale_color_manual(values = c("#FF6666", "#00AFBB", "#E69F00")) + theme_bw() 
```


references:
1. Causal inference for ordinal outcomes https://arxiv.org/pdf/1501.01234.pdf
2. Treatment Effects on Ordinal Outcomes: Causal Estimands and Sharp Bounds https://journals.sagepub.com/doi/pdf/10.3102/1076998618776435
3. drug is not useful https://aasldpubs.onlinelibrary.wiley.com/doi/pdf/10.1002/hep.1840050613







